Long range correlations and their event scale dependence in high multiplicity pp collisions at $\sqrt{s} = 13$~TeV

### 2020-06-12 paper.tex
JL and BK lastet

Long-range azimuthal correlations are measured in very high-multiplicity pp collisions at center-of-mass energy $\sqrt{s} = 13$~TeV by the ALICE experiment at the Large Hadron Collider. 
These results are obtained in the relative pseudorapidity range $1.6 < |\eta| < 1.8$ for the transverse momentum range $1.0 < p_{\mathrm T} < 5.0$~GeV/$c$ and 0--0.1\% multiplicity event class.
In addition to the conventional long-range correlation studies, the results are further investigated in events, where jets with harder fragmentations is to be present.
The jets are reconstructed by using the anti-$k_\mathrm{T}$ algorithm with resolution parameter $R=0.4$ in $|\eta_\mathrm{jet}|<0.4$. 
The magnitude of the correlation dubbed ridge yield, is consistent with the results from other LHC experiments for unbiased events. 
These correlations also persists in the events containing jets. Furthermore data show that the ridge yields tend to increase in events holding $p_{\mathrm T, jet} > 10$~GeV/$c$.
The results are compared to PYTHIA8 string shoving, EPOS and A Multi-Phase Transport ({AMPT}) model calculations.  
PYTHIA8 string shoving model can describe low $p_{\mathrm T}$  ridge yields but can't give good description of  $p_{\mathrm T}$ dependence both for the unbiased and jet events.
The results open new aspects of studying the small collision systems with jet tagged events in the future and will help to explore possible physical origins of long-range correlations.

>>>>

Long-range azimuthal correlations are measured in very high-multiplicity pp collisions at center-of-mass energy $\sqrt{s} = 13$~TeV by ALICE at LHC. 
These results are obtained in a relative pseudorapidity range of $1.6 < |\eta| < 1.8$ and transverse momentum region of $1.0 < p_{\mathrm T} < 5.0$~GeV/$c$ for the trigger and associated particles.
In addition to the conventional long-range correlation studies, the results are further investigated in events, where jets with a harder fragmentation are present.
The jets are reconstructed by using the anti-$k_\mathrm{T}$ algorithm with resolution parameter $R=0.4$ in $|\eta_\mathrm{jet}|<0.4$. 
The magnitude of the correlation dubbed ridge yield, is consistent with the results from other LHC experiments for unbiased events. 
These correlations also persists in the events containing jets. Furthermore data show that the ridge yields tend to increase in events holding $\ptjet > 10$~GeV/$c$.
The results are compared to $\pythiashoving$, EPOS and A Multi-Phase Transport ({AMPT}) model calculations.  
$\pythiashoving$ model can describe low $p_{\mathrm T}$  ridge yields but can't give good description of  $p_{\mathrm T}$ dependence both for the unbiased and jet events.
The results open new aspects of studying the small collision systems with jet tagged events in the future and will help to explore possible physical origins of long-range correlations.

### 2020-06-12 sec_ana.tex
JL and BK lastet

%The multiplicity class is classified by the signal amplitude of the V0 detector.
The high multiplicity events used in the present analysis have the top 0.1\% signal amplitude by the V0 detector with respect to the set of minimum bias events at ALICE. Corresponding number of charged particles is reported as $\sim$31 in the mid-rapidity region ($|\eta|<0.5$) that is almost 5 times larger than the minimum-bias events.  Statistics of the high multiplicity event benefits from a dedicated high multiplicity trigger that was recently implemented at ALICE  in the LHC Run 2 period.

@@

The yield of two-particle correlations between the trigger and associated particles are measured as a function of relative pseudorapidity ($\Delta\eta$) and azimuthal angle ($\Delta\varphi$) differences
\begin{eqnarray}
\frac{1}{N_{\rm{trig}}} \frac{ \rm{d}\it{}^{2} N_{\rm{pair}} }{ \rm{d} \Delta\eta \rm{d}\Delta\varphi} = B(0, 0)\frac{S(\Delta\eta, \Delta\varphi)}{B(\Delta\eta, \Delta\varphi)}\quad,
\end{eqnarray}
where  $p_{\mathrm{T, trig}}$ and $p_{\mathrm{T, assoc}}$ are the transverse momenta of  trigger particles and associated particles, respectively, $N_\mathrm{trig}$ is the number of trigger particles, and $N_\mathrm{pair}$ is the number of pairs of the trigger and associated particles, and $S (\Delta\eta, \Delta\varphi)$ and $B (\Delta\eta, \Delta\varphi)$ are the signal and background distributions, respectively. $S (\Delta\eta, \Delta\varphi)$ is constructed using two-particle correlations in the same event and $B(\Delta\eta, \Delta\varphi)$ is the yield of two-particle correlations in mixed events. The background distribution is normalized by its value at $\Delta\eta$ and $\Delta\varphi = 0$, represented as $B (0,0)$ . In mixed events, the primary vertices and multiplicity classes are required to be in the same range to avoid the acceptance effect with respect to the multiplicity class and primary vertex.

 The quantitative study of ridge is done with $\Delta\varphi$ distribution at a large $\Delta\eta$ to achieve direct comparison of ridges between different event classes and transverse momentum intervals. The large $\Delta\eta$ range can be selected as $1.6<|\Delta\eta|<1.8$ where non-flow effects (mainly coming from jets) are rarely accepted. In this region, jets are not supposed to contribute to the $\Delta\varphi$ distribution that is expressed as
\begin{eqnarray}
\frac{1}{N_{\rm{trig}}} \frac{ \rm{d}\it{}N_{\rm{pair}} }{ \rm{d}\Delta\varphi } = \int_{|\Delta \eta|>1.6} \rm{d} \Delta \eta \frac{1}{\it{N}_{\rm{trig}}} \frac{ \rm{d}\it{}^{2} N_{\rm{pair}} }{ \rm{d}\Delta\eta d\Delta\varphi} \quad.
\end{eqnarray}

The baseline of the correlations is subtracted by implementing Zero-Yield-At-Minimum (ZYAM) procedure. The minimum yield ($C_{\rm{ZYAM}}$) at the minimum $\Delta\varphi$($\Delta\varphi_{\rm{min}}$) of the $\Delta\varphi$ distribution are obtained from the function, which fits the $\Delta\varphi$ distribution with Fourier series up to the third harmonic. Subtracting $C_{\rm{ZYAM}}$ from the $\Delta\varphi$ distribution makes the yield at $\Delta\varphi_{\rm{min}}$ zero in order to describe the shape to find the minimum to be used for integral of the yield. The associated yield of the ridge $Y^{\rm{assoc}}$ is obtained by integrating the near-side peak of the $\Delta\varphi$ distribution over $|\Delta\varphi|<|\Delta\varphi_{\rm{min}}|$ after the ZYAM procedure
\begin{eqnarray}
Y^{\rm{assoc}} = \int_{|\Delta \varphi| < |\Delta\varphi_{\rm{min}}| } \rm{d} \Delta\varphi \frac{1}{\it{N}_{\rm{trig}}} \frac{ \rm{d}\it{}N_{\rm{pair}} }{ \rm{d}\Delta\varphi } \quad.
\end{eqnarray}

The ridge yield is further studied with respect to the magnitude of the hard process. Events with the hard process are selected by requiring minimum transverse momentum of leading track or jet reconstruction in mid-rapidity since leading tracks with a high $p_\mathrm{T}$ or jets result from the deep inelastic scattering of the initial partons. The leading track is selected within $|\Delta\eta|<0.9$. Jets are reconstructed with charged particles only (charged jets) for the anti-$k_{\rm{T}}$ algorithm and the resolution parameter $R = 0.4$. Jets are selected within $|\eta_\mathrm{jet}|<0.4$. The transverse momentum of jets $p_\mathrm{T,\,jet}$ is corrected for underlying event density $rho$.

Monte Carlo simulation with PYTHIA 8 event generator and particle transport inside ALICE using GEANT 3 is used to correct the acceptance and the efficiency of ALICE detectors.

@@

\section{Systematic Uncertainty}
\label{sec:uncertainties}
The systematic uncertainty is evaluated by varying the event selection, track selection, yield extraction procedures and detector efficiency correction. Each source is assumed uncorrelated. The uncertainty from the event selection can be divided into pile-up rejection and primary vertex selection. The uncertainty from the track selection is estimated by changing the selection criteria. The uncertainty from the yield extraction procedures can be categorized into long-range definition and ZYAM procedure. The uncertainty from the efficiency correction is estimated by enlarging the event mixing bins, which is relevant to the acceptance correction.
%Finally, The uncertainty from the application of the detector efficiency correction is estimated comparing the results constructed by efficiency corrected tracks with the results constructed. 
Finally, the uncertainty from the efficiency correction for charged tracks is estimated by comparing the correlations function constructed by reconstructed tracks from the detector responses with the correlations function constructed by generated particles from the event generator.
The uncertainty of $Y^{\rm{assoc}}(\it{p}_{\rm{T, trig}},\,\it{p}_{\rm{T, assoc}})$ is summarized by averaging each uncertainty for given $\it{p}_{\rm{T, trig}}$ and $\it{p}_{\rm{T, assoc}}$. The uncertainty of $Y^{\rm{assoc}}(\it{p}_{\rm{T, Lead}})$ and $Y^{\rm{assoc}}(\it{p}_{\rm{T, Jet}})$ is summarized by averaging the each uncertainty for a given $\it{p}_{\rm{T, Lead}}$ and $\it{p}_{\rm{T, Jet}}$ selection, respectively.
\begin{table}[!h]
\caption{ The relative systematic uncertainty of the associated yield spectrum as function of 
$(\it{p}_{\rm{T, trig}},\,\it{p}_{\rm{T, assoc}})$ (second), $\it{p}_{\rm{T,Lead}}$ (thrid) or $\it{p}_{\rm{T,jet}}$ (fourth) selection in high multiplicity (0--0.1\%) }
\centering
\begin{tabular}{|c|c|c|c|}
\hline 
\multirow{2}{*}{Sources}  & \multicolumn{3}{c|}{Systematic uncertainty error (\%)} \\\cline{2-4}
         & $Y^{\rm{assoc}}(\it{p}_{\rm{T, trig}},\,\it{p}_{\rm{T, assoc}})$ & $Y^{\rm{assoc}}(\it{p}_{\rm{T, Lead}})$ & $Y^{\rm{assoc}}(\it{p}_{\rm{T, Jet}})$ \\ \hline \hline
Pileup cut			& $\sim$0.8	&0-4		&0-4		\\ \hline
Primary Vertex		& $\sim$2.4	&1-12	&1-8		\\ \hline

@@

ZYAM			& $\sim$5.1	&2		&2		\\ \hline
$\Delta\eta$ range	& $\sim$4.5	&3-7		&3-9		\\ \hline

Mixing			& $\sim$4.4	&2-8		&1-16	\\ \hline

M.C. non-closure*	& $\sim$2.5 	&1		&1		\\ \hline
Total 			& $\sim$9.7	&4-16	&4-22	\\ 
\hline 
\end{tabular}

@@

The uncertainty from the pileup rejection is estimated by measuring the changes of results with different rejection criteria from the default one. The pileup rejection is achieved by reconstructing the multiple vertex in the individual event. The multiplicity of the reconstructed vertices determines whether the event is contaminated by pileup or not. The variation is set by the minimum multiplicity of reconstructed vertices from 5 to 3. 
%Pileup cut (multbins) -> (default pileup from the physics selection)
%Pileup cut (multbins) -> PileupMV (?)
The estimated uncertainty is 1\% without $\it{p}_{\rm{T, Lead}}$ or $\it{p}_{\rm{T, Jet}}$ selection. The uncertainty is increased up to 4\% with $\it{p}_{\rm{T, Lead}}$ or $\it{p}_{\rm{T, Jet}}$ selection. The uncertainty from the primary vertex selection along the beam axis is estimated by varying the selection range of primary vertex from $|z_\mathrm{vtx}|<$ 8 cm to $|z_\mathrm{vtx}|<$ 6 cm. The narrower primary vertex selection allows one to accept more forward tracks, which could be very sensitive to select long-range side. The estimated uncertainty is increased up to 12\% from 1\%, which is obtained without the event scale selection.
The uncertainty from the track selection is estimated by varying the track selection criteria from the one optimized for refining the angular track distribution to the other optimized for particle identification. The variation is achieved by changing the mendatory SPD requirement and the maximal threshold for chi-square of the track fit.
%https://github.com/alisw/AliRoot/blob/master/ANALYSIS/ESDfilter/macros/AddTaskESDFilter.C
The estimated uncertainty for track selection is 4\% by averaging the uncertainty for whole transverse momentum bins and 2\% for $1<\it{p}_{\rm{T}}<2$ GeV/$c$ tracks.

The uncertainty from the ZYAM procedure is estimated by varying the finding-scope for minimum position. The estimated uncertainty for track selection is 5.1\% by averaging the uncertainty for whole transverse momentum bins and 2\% for 1$<\it{p}_{\rm{T}}<$2 GeV/\it{c}\rm{} tracks, which is not largely affected by the event scale selection. The uncertainty from the long $\Delta\eta$ range is estimated by varying the $\Delta\eta$ range. The uncertainty might include effects from the longitudinal de-correlations and possible non-flow contamination, which is thought be small. The estimated uncertainty is increased up to 9\% from 3\%, which is obtained by averaging the whole transverse momentum range.
The uncertainty from the mixing pool bin size of primary vertex is estimated by varying the bin size from 2.0 cm to 1.0 cm. The estimated uncertainty is 1-16\%

>>>>

%The multiplicity class is classified by the signal amplitude of the V0 detector.
The high multiplicity events used in the present analysis have the top 0.1\%  signal amplitude (the 0--0.1\% multiplicity class) by the V0 detector with respect to the set of minimum bias events at ALICE. Corresponding number of charged particles is reported as $\sim$31 in the mid-rapidity region ($|\eta|<0.5$) that is almost 5 times larger than the minimum-bias events.  Statistics of the high multiplicity event benefits from a dedicated high multiplicity trigger that was recently implemented at ALICE  in the LHC Run 2 period.

@@

The yield of two-particle correlations between the trigger and associated particles are measured as functions of relative pseudorapidity ($\Delta\eta$) and azimuthal angle ($\Delta\varphi$) differences
\begin{eqnarray}
\frac{1}{N_{\rm{trig}}} \frac{ \rm{d}\it{}^{2} N_{\rm{pair}} }{ \rm{d} \Delta\eta \rm{d}\Delta\varphi} = B(0, 0)\frac{S(\Delta\eta, \Delta\varphi)}{B(\Delta\eta, \Delta\varphi)}\quad,
\label{eq:corrfunction}
\end{eqnarray}
where  $\pttrig$ and $\ptassoc$ are the transverse momenta of  the trigger and associated particles, respectively, $N_\mathrm{trig}$ is the number of trigger particles, $N_\mathrm{pair}$ is the number of pairs of the trigger and associated particles, and $S (\Delta\eta, \Delta\varphi)$ and $B (\Delta\eta, \Delta\varphi)$ are the signal and background distributions, respectively. $S (\Delta\eta, \Delta\varphi)$ is constructed using two-particle correlations in the same event and $B(\Delta\eta, \Delta\varphi)$ is the yield of two-particle correlations in mixed events. The background distribution is normalized by its value at $\Delta\eta$ and $\Delta\varphi = 0$, represented as $B (0,0)$. In mixed events, the primary vertices and multiplicity classes are required to be in the same range to avoid the acceptance effect with respect to the multiplicity class and primary vertex.

 The quantitative study of the ridge is done with $\Delta\varphi$ distribution at a large $\Delta\eta$ to achieve direct comparison of ridges between different event classes and transverse momentum intervals. The large $\Delta\eta$ range can be selected as $1.6<|\Delta\eta|<1.8$ where non-flow effects (mainly coming from jets) are rarely accepted. In this region, jets are supposed not to contribute to the $\Delta\varphi$ distribution that is expressed as
\begin{eqnarray}
\frac{1}{N_{\rm{trig}}} \frac{ \rm{d}\it{}N_{\rm{pair}} }{ \rm{d}\Delta\varphi } = \int_{|\Delta \eta|>1.6} \rm{d} \Delta \eta \frac{1}{\it{N}_{\rm{trig}}} \frac{ \rm{d}\it{}^{2} N_{\rm{pair}} }{ \rm{d}\Delta\eta d\Delta\varphi} \quad.
\end{eqnarray}

The baseline of the correlations is subtracted by implementing Zero-Yield-At-Minimum (ZYAM) procedure~\cite{Ajitanand:2005jj}. The minimum yield $(C_{\rm{ZYAM}})$ at the $\Delta\varphi$ ($\Delta\varphi_{\rm{min}}$) in the $\Delta\varphi$ distribution is obtained from the function, which fits the $\Delta\varphi$ distribution with Fourier series up to the third harmonic. Subtracting $C_{\rm{ZYAM}}$ from the $\Delta\varphi$ distribution makes the yield at $\Delta\varphi_{\rm{min}}$ zero in order to describe the shape to find the minimum to be used for integral of the yield. The associated yield of the ridge $Y^{\rm{assoc}}$ is obtained by integrating the near-side peak of the $\Delta\varphi$ distribution over $|\Delta\varphi|<|\Delta\varphi_{\rm{min}}|$ after the ZYAM procedure
\begin{eqnarray}
Y^{\rm{assoc}} = \int_{|\Delta \varphi| < |\Delta\varphi_{\rm{min}}| } \rm{d} \Delta\varphi \frac{1}{\it{N}_{\rm{trig}}} \frac{ \rm{d}\it{}N_{\rm{pair}} }{ \rm{d}\Delta\varphi } \quad.
\end{eqnarray}

The ridge yield is measured with respect to the multiplicity classes. Multiplicity classes are defined by a probability range with respect to the minimum bias events, constituting 100\%. Percentile values are close to 0  for higher multiplicity while close to 100 for lower multiplicity collisions like the centrality classification in heavy-ion collisions~\cite{Abelev:2013qoq,ALICE-PUBLIC-2018-003}. Multiplicity classes are estimated by the V0 detector.The high multiplicity pp collisions in this analysis is defined as the top 0.1\% signal amplitude events measured by the V0 detector. The corresponding multiplicity class is called the 0--0.1\%.

The ridge yield is further studied with respect to the magnitude of the hard process. Events with the hard process are selected by requiring minimum transverse momentum of leading track or jet reconstruction in mid-rapidity since leading tracks with a high $p_\mathrm{T}$ or jets result from the deep inelastic scattering of the initial partons. The leading track is selected within $|\Delta\eta|<0.9$ and the full azimuthal angle. Jets are reconstructed with charged particles only (charged jets) for the anti-$k_{\rm{T}}$ algorithm and the resolution parameter $R = 0.4$. Jets are selected in $|\eta_\mathrm{jet}|<0.4$ and the full azimuthal angle. The transverse momentum of jets $p_\mathrm{T,\,jet}$ is corrected for underlying event density that is measured for the $k_{\rm{T}}$ algorithm with $R = 0.2$.

Monte Carlo simulation with $\pythiam$ event generator and particle transport inside ALICE using GEANT 3 is used to correct the acceptance and the efficiency of the ALICE detectors.

@@

\section{Systematic Uncertainty}
\label{sec:uncertainties}
The systematic uncertainty is evaluated by varying the event selection, track selection, yield extraction procedures and detector efficiency correction. Each source is assumed uncorrelated. The uncertainty from the event selection can be divided into pile-up rejection and primary vertex selection. The uncertainty from the track selection is estimated by changing the selection criteria. The uncertainty from the yield extraction procedures can be categorized into long-range definition and ZYAM procedure. The uncertainty from the efficiency correction is estimated by enlarging the event mixing bins, which is relevant to the acceptance correction.
%Finally, The uncertainty from the application of the detector efficiency correction is estimated comparing the results constructed by efficiency corrected tracks with the results constructed. 
Finally, the uncertainty from the efficiency correction for charged tracks is estimated by comparing the correlations function constructed by reconstructed tracks from the detector responses with the correlations function constructed by generated particles from the event generator.
The uncertainty of $Y^{\rm{assoc}}(\pttrig,\,\ptassoc)$ is summarized by averaging each uncertainty for given $\pttrig$ and $\ptassoc$. The uncertainty of $Y^{\rm{assoc}}(\ptlead)$ and $Y^{\rm{assoc}}(\ptjet)$ is summarized by averaging the each uncertainty for a given $\ptlead$ and $\ptjet$ selection, respectively.
\begin{table}[!h]
\caption{ The relative systematic uncertainty of the associated yield spectrum are estimated for $Y^{\rm{assoc}}(\pttrig,\,\ptassoc)$, $Y^{\rm{assoc}}(\ptlead)$, and $Y^{\rm{assoc}}(\ptjet)$, respectively.  }
\centering
\begin{tabular}{|c|c|c|c|}
\hline 
\multirow{2}{*}{Sources}  & \multicolumn{3}{c|}{Systematic uncertainty (\%)} \\\cline{2-4}
         & $Y^{\rm{assoc}}(\it{p}_{\rm{T, trig}},\,\it{p}_{\rm{T, assoc}})$ & $Y^{\rm{assoc}}(\it{p}_{\rm{T, Lead}})$ & $Y^{\rm{assoc}}(\it{p}_{\rm{T, Jet}})$ \\ \hline \hline
Pileup cut			& $\sim$0.8	&0--4		&0--4		\\ \hline
Primary Vertex		& $\sim$2.4	&1--12	&1-8		\\ \hline

@@

ZYAM			& $\sim$5.1	&2		&2		\\ \hline
$\Delta\eta$ range	& $\sim$4.5	&3-7		&3--9		\\ \hline

Mixing			& $\sim$4.4	&2--8		&1--16	\\ \hline

M.C. non-closure*	& $\sim$2.5 	&1		&1		\\ \hline
Total 			& $\sim$9.7	&4--16	&4--22	\\ 
\hline 
\end{tabular}

@@

The uncertainty from the pileup rejection is estimated by measuring the changes of results with different rejection criteria from the default one. The pileup rejection is achieved by reconstructing the multiple vertex in the individual event. The multiplicity of the reconstructed vertices determines whether the event is contaminated by pileup or not. The variation is set by the minimum multiplicity of reconstructed vertices from 5 to 3. 
%Pileup cut (multbins) -> (default pileup from the physics selection)
%Pileup cut (multbins) -> PileupMV (?)
The estimated uncertainty is 1\% without $\ptlead$ or $\ptjet$ selection. The uncertainty is increased up to 4\% with $\ptlead$ or $\ptjet$ selection. The uncertainty from the primary vertex selection along the beam axis is estimated by varying the selection range of primary vertex from $|z_\mathrm{vtx}|<$ 8 to 6 cm. The narrower primary vertex selection allows one to accept more forward tracks, which could be very sensitive to select the long-range side. The estimated uncertainty is increased up to 12\% from 1\%, which is obtained without the event scale selection.
The uncertainty from the track selection is estimated by varying the track selection criteria from the one optimized for refining the angular track distribution to the other optimized for particle identification. The variation is achieved by changing the mandatory SPD requirement and the maximal threshold for chi-square of the track fit.
%https://github.com/alisw/AliRoot/blob/master/ANALYSIS/ESDfilter/macros/AddTaskESDFilter.C
The estimated uncertainty for track selection is 4\% by averaging the uncertainty for whole transverse momentum bins and 2\% for $1<\it{p}_{\rm{T}}<2$ GeV/$c$ tracks.

The uncertainty from the ZYAM procedure is estimated by varying the finding-scope for minimum position. The estimated uncertainty for track selection is 5.1\% by averaging the uncertainty for the whole transverse momentum bins and 2\% for $1<\it{p}_{\rm{T}}<2$ GeV/$c$ tracks, which is not largely affected by the event scale selection. The uncertainty from the long $\Delta\eta$ range is estimated by varying the $\Delta\eta$ range. The uncertainty might include effects from the longitudinal de-correlations and possible non-flow contamination, which is thought be small. The estimated uncertainty is increased up to 9\% from 3\%, which is obtained by averaging the whole transverse momentum range.
The uncertainty from the mixing pool bin size of primary vertex is estimated by varying the bin size from 2 to 1 cm. The estimated uncertainty is 1--16\%
